---
title: "Relatedness: YOY Blacktips in US. Waters"
author: "DG Swift"
output:
  html_document:
    df_print: paged
    toc: yes
  html_notebook:
    toc: yes
---
  
# Environment
  
```{r, message=FALSE}

# set working directory

.libPaths("/usr/lib64/R/library")

# invalidate cache when the package version changes

knitr::opts_chunk$set(
  root.dir = "~/Projects/us_blacktips/related",
	message = FALSE,
	warning = FALSE,
	cache.extra = packageVersion("tint"),
	tidy = FALSE,
	echo = FALSE)

options(htmltools.dir.version = FALSE)

# conflicts

library(conflicted)
conflict_prefer("count", "dplyr")
conflict_prefer("arrange", "dplyr")
conflict_prefer("mutate", "dplyr")
conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")
conflict_prefer("rename", "dplyr")
conflict_prefer("summarise", "dplyr")
conflict_prefer("summarize", "dplyr")
conflict_prefer("s.label", "adegraphics")
conflict_prefer("s.value", "adegraphics")
conflict_prefer("scalebar", "raster")
conflict_prefer("rename", "dplyr")

# packages

library(tidyverse)
library(broom)
library(adegenet)
library(hierfstat)
library(pegas)
library(Demerelate)
library(ggthemes)
library(related)
library(poppr)
library(glue)
library(janitor)
library(here)
library(readxl)
library(zvau)
library(patchwork)

# source functions

source("~/code/genind.R")
source("~/code/PCA.R")
source("~/code/ggplot.R")
source("~/code/DAPC.R")

`%not in%` <- function (x, table) is.na(match(x, table, nomatch=NA_integer_))

# orders, colors, shapes

site_order_shrt <- c("BLB", "SHS", "PRS", "TCB", "WAB", "APB", "MOB", "GAB", "MAB", "SAB", "CCB")

region_col <- c('#1b9e77','#d95f02','#7570b3')

region_shape <- c(21, 23, 24)

site_shape <- c(21, 21, 21, 23, 23, 23, 23, 24, 24, 24, 24)

relationship_order <- c("unrelated", "halfsib", "fullsib")

col_rel <- c("#969696", "#525252", "#252525")

```

# Import Data

```{r, message=F}

# genepop

yoy.gen <- read.genepop(file = here("filter", "results", "yoy_final_site.gen"), ncode = 3L, quiet = FALSE)

yoy_inds <- as_tibble_col(indNames(yoy.gen), column_name = "hiseq_id")

# produce strata

yoy_strata <- read_csv(here("filter", "results", "yoy_final_strata.csv")) %>%
  left_join(yoy_inds, .)

# assign strata

strata(yoy.gen) <- yoy_strata

setPop(yoy.gen) <- ~site_shrt

```

# All Individuals

In a separate R terminal, run Wang's `trioml` method for all individuals to produce point estimates.

Before running, remove loci with any missing data to reduce the run time. 

```{r warning=FALSE, message=FALSE, eval = FALSE}

library(related)
library(adegenet)
library(tidyverse)
library(poppr)
library(hierfstat)
library(here)

# genepop

yoy.gen <- read.genepop(file = here("filter", "results", "yoy_final_site.gen"), ncode = 3L, quiet = FALSE)

yoy_inds <- as_tibble_col(indNames(yoy.gen), column_name = "hiseq_id")

# produce strata

yoy_strata <- read_csv(here("filter", "results", "yoy_final_strata.csv")) %>%
  left_join(yoy_inds, .)

# assign strata

strata(yoy.gen) <- yoy_strata

setPop(yoy.gen) <- ~site_shrt

# remove loci with missing data

temp.gen <- missingno(yoy.gen, type = "loci", cutoff = 0)

## 2230 loci remaining ##

# point estimates

temp_df <- genind2df(temp.gen, usepop = FALSE, oneColPerAll = TRUE) %>%
  rownames_to_column("hiseq_id")

write_delim(temp_df, here("related", "data", "trioml.input"), col_names = FALSE, delim = "\t")

genotypedata <- readgenotypedata(here("related", "data", "trioml.input"))

related <- coancestry(genotypedata$gdata, trioml = 1)

related <- related$relatedness

write_csv(related, here("related", "data", "trioml_point.output"))

```

Redo to produce confidence intervals. 

```{r warning=FALSE, message=FALSE, eval = FALSE}

library(related)
library(adegenet)
library(tidyverse)
library(poppr)
library(hierfstat)
library(here)

# genepop

yoy.gen <- read.genepop(file = here("filter", "results", "yoy_final_site.gen"), ncode = 3L, quiet = FALSE)

yoy_inds <- as_tibble_col(indNames(yoy.gen), column_name = "hiseq_id")

# produce strata

yoy_strata <- read_csv(here("filter", "results", "yoy_final_strata.csv")) %>%
  left_join(yoy_inds, .)

# assign strata

strata(yoy.gen) <- yoy_strata

setPop(yoy.gen) <- ~site_shrt

# remove loci with missing data

temp.gen <- missingno(yoy.gen, type = "loci", cutoff = 0)

## 2230 loci remaining ##

# point estimates

temp_df <- genind2df(temp.gen, usepop = FALSE, oneColPerAll = TRUE) %>%
  rownames_to_column("hiseq_id")

write_delim(temp_df, here("related", "data", "trioml.input"), col_names = FALSE, delim = "\t")

genotypedata <- readgenotypedata(here("related", "data", "trioml.input"))

related <- coancestry(genotypedata$gdata, trioml = 2)

related_output <- related$relatedness.ci95

write_csv(related_output, here("related", "data", "trioml_ci.output"))

```

## Results

Import results and use threshold hold of 0.2 to assess putative siblings.

```{r}

# point estimates

trioml_point <- read_csv(here("related", "data", "trioml_point.output")) %>% 
  select(1:5)

temp1 <- trioml_point %>%
  select(pair.no, ind1.id, ind2.id, trioml) %>%
  rename(point = trioml) %>%
  rename(hiseq_id = ind1.id) %>%
  left_join(yoy_strata) %>%
  select(c(hiseq_id, sample_id, stage, sex, site_shrt, region, year, month,  point))
  
temp2 <- trioml_point %>%
  select(ind2.id) %>%
  rename(hiseq_id = ind2.id) %>%
  left_join(yoy_strata) %>%
  select(c(hiseq_id, sample_id, stage, sex, site_shrt, region, year, month)) %>%
  rename(hiseq_id_2 = hiseq_id) %>%
  rename(sample_id_2 = sample_id) %>%
  rename(stage_2 = stage) %>%
  rename(sex_2 = sex) %>%
  rename(site_shrt_2 = site_shrt) %>%
  rename(region_2 = region) %>%
  rename(year_2 = year) %>%
  rename(month_2 = month)

# combine

trioml_point <- cbind(temp1, temp2)

# ci estimates

trioml_ci <- read_csv(here("related", "data", "trioml_ci.output")) %>% 
  select("trioml.low", "trioml.high") %>% 
  rename(low = trioml.low) %>%
  rename(high = trioml.high)

# combine

trioml_results <- cbind(trioml_point, trioml_ci) %>% 
  relocate("low", .before="point") %>% 
  relocate("high", .after="point") %>% 
  mutate(across(c(low, point, high), round, 2)) %>% 
  arrange(desc(point))

# export results

write_csv(trioml_results, here("related", "results", "yoy_trioml_related.csv"))

# putative siblings

trioml_sibs <- trioml_results %>% 
  filter(high >= 0.25)

write_csv(trioml_sibs, here("related", "results", "yoy_trioml_sibs.csv"))

count(trioml_sibs %>%
  filter("region" == "region_2"))

count(trioml_sibs %>%
  filter("site_shrt" == "site_shrt_2"))

```

# By Region

For the Atlantic, eastern and western Gulf individually, assess relatedness using `demerelate`.

Atlantic.

```{r warning=FALSE, message=FALSE, eval = FALSE}

library(Demerelate)
library(adegenet)
library(tidyverse)
library(poppr)
library(hierfstat)
library(here)
source("~/code/genind.R")

# genepop

yoy.gen <- read.genepop(file = here("filter", "results", "yoy_final_site.gen"), ncode = 3L, quiet = FALSE)

yoy_inds <- as_tibble_col(indNames(yoy.gen), column_name = "hiseq_id")

# produce strata

yoy_strata <- read_csv(here("filter", "results", "yoy_final_strata.csv")) %>%
  left_join(yoy_inds, .)

# execute demerelate for each region

## Atl

atl_strata <- yoy_strata %>%
  filter(region == "Atl")

remove_inds <- anti_join(yoy_strata, atl_strata)
remove_inds <- remove_inds$hiseq_id

atl.gen <- gen.ind.rem.Ind(yoy.gen, remove_inds)

atl_inds <- as_tibble_col(indNames(atl.gen), column_name = "hiseq_id")

atl_strata <- left_join(atl_inds, atl_strata)

strata(atl.gen) <- atl_strata

setPop(atl.gen) <- ~region

# generate genetic diversity stats

gendiv <- adegenet::summary(atl.gen)

# remove monomorphic loci

ho <- as.data.frame(gendiv$Hobs) %>% 
  rownames_to_column("LOCUS") %>%
  rename(ho = `gendiv$Hobs`)

monomorphic <- filter(ho, ho == 0)

remove_loci <- monomorphic$LOCUS

atl.gen <- genind.rem.loci(atl.gen, remove_loci)

atl.gen <- missingno(atl.gen, type = "loci", cutoff = 0)

atl_df <- genind2df(atl.gen, usepop = TRUE, oneColPerAll = TRUE) %>%
  rownames_to_column("hiseq_id")

demerelate <- Demerelate(atl_df, tab.dist = FALSE, ref.pop = FALSE, object=TRUE, value="wang", Fis = FALSE, iteration = 1000, pairs = 1000, file.output = TRUE, p.correct = TRUE, NA.rm = TRUE, genotype.ref = FALSE)

```

Eastern Gulf

```{r warning=FALSE, message=FALSE, eval = FALSE}

library(Demerelate)
library(adegenet)
library(tidyverse)
library(poppr)
library(hierfstat)
library(here)
source("~/code/genind.R")

# genepop

yoy.gen <- read.genepop(file = here("filter", "results", "yoy_final_site.gen"), ncode = 3L, quiet = FALSE)

yoy_inds <- as_tibble_col(indNames(yoy.gen), column_name = "hiseq_id")

# produce strata

yoy_strata <- read_csv(here("filter", "results", "yoy_final_strata.csv")) %>%
  left_join(yoy_inds, .)

# execute demerelate for each region

## EGoM

egom_strata <- yoy_strata %>%
  filter(region == "EGoM")

remove_inds <- anti_join(yoy_strata, egom_strata)
remove_inds <- remove_inds$hiseq_id

egom.gen <- gen.ind.rem.Ind(yoy.gen, remove_inds)

egom_inds <- as_tibble_col(indNames(egom.gen), column_name = "hiseq_id")

egom_strata <- left_join(egom_inds, egom_strata)

strata(egom.gen) <- egom_strata

setPop(egom.gen) <- ~region

# generate genetic diversity stats

gendiv <- adegenet::summary(egom.gen)

# remove monomorphic loci

ho <- as.data.frame(gendiv$Hobs) %>% 
  rownames_to_column("LOCUS") %>%
  rename(ho = `gendiv$Hobs`)

monomorphic <- filter(ho, ho == 0)

remove_loci <- monomorphic$LOCUS

egom.gen <- genind.rem.loci(egom.gen, remove_loci)

egom.gen <- missingno(egom.gen, type = "loci", cutoff = 0)

egom_df <- genind2df(egom.gen, usepop = TRUE, oneColPerAll = TRUE) %>%
  rownames_to_column("hiseq_id")

demerelate <- Demerelate(egom_df, tab.dist = FALSE, ref.pop = FALSE, object=TRUE, value="wang", Fis = FALSE, iteration = 1000, pairs = 1000, file.output = TRUE, p.correct = TRUE, NA.rm = TRUE, genotype.ref = FALSE)

```

Western Gulf

```{r warning=FALSE, message=FALSE, eval = FALSE}

library(Demerelate)
library(adegenet)
library(tidyverse)
library(poppr)
library(hierfstat)
library(here)
source("~/code/genind.R")

# genepop

yoy.gen <- read.genepop(file = here("filter", "results", "yoy_final_site.gen"), ncode = 3L, quiet = FALSE)

yoy_inds <- as_tibble_col(indNames(yoy.gen), column_name = "hiseq_id")

# produce strata

yoy_strata <- read_csv(here("filter", "results", "yoy_final_strata.csv")) %>%
  left_join(yoy_inds, .)

# execute demerelate for each region

## WGoM

wgom_strata <- yoy_strata %>%
  filter(region == "WGoM")

remove_inds <- anti_join(yoy_strata, wgom_strata)
remove_inds <- remove_inds$hiseq_id

wgom.gen <- gen.ind.rem.Ind(yoy.gen, remove_inds)

wgom_inds <- as_tibble_col(indNames(wgom.gen), column_name = "hiseq_id")

wgom_strata <- left_join(wgom_inds, wgom_strata)

strata(wgom.gen) <- wgom_strata

setPop(wgom.gen) <- ~region

# generate genetic diversity stats

gendiv <- adegenet::summary(wgom.gen)

# remove monomorphic loci

ho <- as.data.frame(gendiv$Hobs) %>% 
  rownames_to_column("LOCUS") %>%
  rename(ho = `gendiv$Hobs`)

monomorphic <- filter(ho, ho == 0)

remove_loci <- monomorphic$LOCUS

wgom.gen <- genind.rem.loci(wgom.gen, remove_loci)

wgom.gen <- missingno(wgom.gen, type = "loci", cutoff = 0)

wgom_df <- genind2df(wgom.gen, usepop = TRUE, oneColPerAll = TRUE) %>%
  rownames_to_column("hiseq_id")

demerelate <- Demerelate(wgom_df, tab.dist = FALSE, ref.pop = FALSE, object=TRUE, value="wang", Fis = FALSE, iteration = 1000, pairs = 1000, file.output = TRUE, p.correct = TRUE, NA.rm = TRUE, genotype.ref = FALSE)

```

Move `demerelate` directory from to `relatedness` directory. The name of the `demerelate` directory changes with each run.

```{bash}

# you may have to look for the demerelate folders and change the path

cd ~/Projects/us_blacktips/filter/data/

mv Demerelate* ~/Projects/us_blacktips/related/data/.

```

## Results

#### Atlantic

`Demerelate` uses simulations to generate pairs of full and half siblings and unrelated individuals. Use these relationships to determine the thresholds for identifying full and half siblings. 

```{r fig.height=5, fig.width=5, message=FALSE, warning=FALSE}

sim_ur_atl <- read.table(here("related", "data", "Demerelate-0.2835", "Random.NonRelated.relatedness.overall.txt")) %>%
  rename(unrelated = wang)

sim_hs_atl <- read.table(here("related", "data", "Demerelate-0.2835", "Random.Halfsib.relatedness.overall.txt")) %>%
  rename(halfsib = wang)

sim_fs_atl <- read.table(here("related", "data", "Demerelate-0.2835", "Random.Fullsib.relatedness.overall.txt")) %>%
  rename(fullsib = wang)

# combine

sim_atl <- cbind(sim_ur_atl, sim_hs_atl, sim_fs_atl) %>% 
  gather(relationship, wang) %>%
  mutate(relationship = ordered(relationship, levels = relationship_order)) %>% 
  mutate_if(is.numeric, round, 2)

# plot histograms of relatedness 

atl_plot <- ggplot(sim_atl, aes(x=wang, fill = relationship, label = relationship, color = relationship)) +
  geom_histogram(binwidth = 0.01, alpha=0.6, position = 'identity', show.legend = F) +
  #geom_vline(aes(xintercept = 0),
   #          color = "black", linetype = "solid", size = 0.75) +
  geom_vline(aes(xintercept = 0.25),
             color = "black", linetype = "solid", size = 0.75) +
  geom_vline(aes(xintercept = 0.5),
             color = "black", linetype = "solid", size = 0.75) +
  geom_vline(aes(xintercept = quantile(sim_hs_atl, 0.01, na.rm = TRUE)),
             color = "red", linetype = "dotted", size = 0.75) +
  geom_vline(aes(xintercept = quantile(sim_fs_atl, 0.01, na.rm = TRUE)),
             color = "red", linetype = "dotted", size = 0.75) +
  labs(x = "Relatedness", y = "Number of Pairs") +
  scale_fill_manual(values=col_rel) +
  scale_color_manual(values=col_rel) +
  ggtitle("Atlantic") +
  theme_standard + 
  theme(plot.title = element_text(hjust = 0.5))

atl_plot

ggsave(here("related", "results", "atl_related_sim.png"))

# full sib threshold

fs_thres_atl <- round(quantile(sim_fs_atl, 0.01, na.rm = TRUE), 2)
fs_thres_atl

# half sib threshold

hs_thres_atl <- round(quantile(sim_hs_atl, 0.01, na.rm = TRUE), 2)
hs_thres_atl

```

Identify siblings.

```{r fig.height=5, fig.width=5, message=FALSE, warning=FALSE}

atl_related <- read.table(here("related", "data", "Demerelate-0.2835", "Empirical.relatedness.Atl.txt")) %>%
  rownames_to_column() %>%
  separate(rowname, into = c("ind1", "ind2"), sep = "_(?=[A-Z])", remove = TRUE) %>%
  mutate_if(is.numeric, round, 5) %>%
  mutate(pop = "Atl")

# match to strata

related1 <- atl_related %>%
  select(ind1, ind2, wang) %>%
  rename(hiseq_id = ind1) %>%
  left_join(yoy_strata) %>%
  select(c(hiseq_id, sample_id, stage, sex, region, site_shrt, latitude, longitude, year, month, tl, wang))

related2 <- atl_related %>%
  select(ind2) %>%
  rename(hiseq_id = ind2) %>%
  left_join(yoy_strata) %>%
  select(c(hiseq_id, sample_id, stage, sex, region, site_shrt, latitude, longitude, year, month, tl)) %>%
  rename(hiseq_id_2 = hiseq_id) %>%
  rename(sample_id_2 = sample_id) %>%
  rename(stage_2 = stage) %>%
  rename(sex_2 = sex) %>%
  rename(region_2 = region) %>%
  rename(site_shrt_2 = site_shrt) %>%
  rename(latitude_2 = latitude) %>%
  rename(longitude_2 = longitude) %>%
  rename(year_2 = year) %>%
  rename(month_2 = month) %>%
  rename(tl_2 = tl)

atl_related <- cbind(related1, related2) %>%
  arrange(desc(wang)) %>% 
  mutate_at("wang", round, 2)

write_csv(atl_related, here("related", "results", "atl_related.csv"))

# Plot

ggplot(atl_related, aes(x = wang)) +
  geom_histogram(binwidth = 0.01, color = "black", fill = "grey") +
  geom_vline(aes(xintercept = mean(wang)),
             color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0),
             color = "black", linetype = "solid", size = 1) +
  labs(x = "Relatedness", y = "Number of Pairs") +
  ggtitle("Atlantic") +
  theme_standard + 
  theme(plot.title = element_text(hjust = 0.5))

ggsave(here("related", "results", "atl_related.png"))

# full siblings

fs_atl <- atl_related %>% 
  filter(wang >= fs_thres_atl)

# half siblings

hs_atl <- atl_related %>% 
  filter(wang >= hs_thres_atl & wang <= fs_thres_atl)

```

#### Eastern Gulf

```{r fig.height=5, fig.width=5, message=FALSE, warning=FALSE}

sim_ur_egom <- read.table(here("related", "data", "Demerelate-0.1564", "Random.NonRelated.relatedness.overall.txt")) %>%
  rename(unrelated = wang)

sim_hs_egom <- read.table(here("related", "data", "Demerelate-0.1564", "Random.Halfsib.relatedness.overall.txt")) %>%
  rename(halfsib = wang)

sim_fs_egom <- read.table(here("related", "data", "Demerelate-0.1564", "Random.Fullsib.relatedness.overall.txt")) %>%
  rename(fullsib = wang)

# combine

sim_egom <- cbind(sim_ur_egom, sim_hs_egom, sim_fs_egom) %>% 
  gather(relationship, wang) %>%
  mutate(relationship = ordered(relationship, levels = relationship_order)) %>% 
  mutate_if(is.numeric, round, 2)

# plot histograms of relatedness 

egom_plot <- ggplot(sim_egom, aes(x=wang, fill = relationship, label = relationship, color = relationship)) +
  geom_histogram(binwidth = 0.01, alpha=0.6, position = 'identity') +
#  geom_vline(aes(xintercept = 0),
 #            color = "black", linetype = "solid", size = 0.75) +
  geom_vline(aes(xintercept = 0.25),
             color = "black", linetype = "solid", size = 0.75) +
  geom_vline(aes(xintercept = 0.5),
             color = "black", linetype = "solid", size = 0.75) +
  geom_vline(aes(xintercept = quantile(sim_hs_egom, 0.01, na.rm = TRUE)),
             color = "red", linetype = "dotted", size = 0.75) +
  geom_vline(aes(xintercept = quantile(sim_fs_egom, 0.01, na.rm = TRUE)),
             color = "red", linetype = "dotted", size = 0.75) +
  labs(x = "Relatedness", y = "Number of Pairs") +
  scale_fill_manual(values=col_rel) +
  scale_color_manual(values=col_rel) +
  ggtitle("Eastern Gulf") +
  theme_standard + 
  theme(plot.title = element_text(hjust = 0.5))

egom_plot

ggsave(here("related", "results", "egom_related_sim.png"))

# full sib threshold

fs_thres_egom <- round(quantile(sim_fs_egom, 0.01, na.rm = TRUE), 2)
fs_thres_egom

# half sib threshold

hs_thres_egom <- round(quantile(sim_hs_egom, 0.01, na.rm = TRUE), 2)
hs_thres_egom

```

Identify siblings.

```{r fig.height=5, fig.width=5, message=FALSE, warning=FALSE}

egom_related <- read.table(here("related", "data", "Demerelate-0.1564", "Empirical.relatedness.EGoM.txt")) %>%
  rownames_to_column() %>%
  separate(rowname, into = c("ind1", "ind2"), sep = "_(?=[A-Z])", remove = TRUE) %>%
  mutate_if(is.numeric, round, 5) %>%
  mutate(pop = "egom")

# match to strata

related1 <- egom_related %>%
  select(ind1, ind2, wang) %>%
  rename(hiseq_id = ind1) %>%
  left_join(yoy_strata) %>%
  select(c(hiseq_id, sample_id, stage, sex, region, site_shrt, latitude, longitude, year, month, tl, wang))

related2 <- egom_related %>%
  select(ind2) %>%
  rename(hiseq_id = ind2) %>%
  left_join(yoy_strata) %>%
  select(c(hiseq_id, sample_id, stage, sex, region, site_shrt, latitude, longitude, year, month, tl)) %>%
  rename(hiseq_id_2 = hiseq_id) %>%
  rename(sample_id_2 = sample_id) %>%
  rename(stage_2 = stage) %>%
  rename(sex_2 = sex) %>%
  rename(region_2 = region) %>%
  rename(site_shrt_2 = site_shrt) %>%
  rename(latitude_2 = latitude) %>%
  rename(longitude_2 = longitude) %>%
  rename(year_2 = year) %>%
  rename(month_2 = month) %>%
  rename(tl_2 = tl)

egom_related <- cbind(related1, related2) %>%
  arrange(desc(wang)) %>% 
  mutate_at("wang", round, 2)

write_csv(egom_related, here("related", "results", "egom_related.csv"))

# Plot

ggplot(egom_related, aes(x = wang)) +
  geom_histogram(binwidth = 0.01, color = "black", fill = "grey") +
  geom_vline(aes(xintercept = mean(wang)),
             color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0),
             color = "black", linetype = "solid", size = 1) +
  labs(x = "Relatedness", y = "Number of Pairs") +
  ggtitle("Eastern Gulf") +
  theme_standard + 
  theme(plot.title = element_text(hjust = 0.5))

ggsave(here("related", "results", "egom_related.png"))

# full siblings

fs_egom <- egom_related %>% 
  filter(wang >= fs_thres_egom)

# half siblings

hs_egom <- egom_related %>% 
  filter(wang >= hs_thres_egom & wang <= fs_thres_egom)

```

#### Western Gulf

```{r fig.height=5, fig.width=5, message=FALSE, warning=FALSE}

sim_ur_wgom <- read.table(here("related", "data", "Demerelate0.7719", "Random.NonRelated.relatedness.overall.txt")) %>%
  rename(unrelated = wang)

sim_hs_wgom <- read.table(here("related", "data", "Demerelate0.7719", "Random.Halfsib.relatedness.overall.txt")) %>%
  rename(halfsib = wang)

sim_fs_wgom <- read.table(here("related", "data", "Demerelate0.7719", "Random.Fullsib.relatedness.overall.txt")) %>%
  rename(fullsib = wang)

# combine

sim_wgom <- cbind(sim_ur_wgom, sim_hs_wgom, sim_fs_wgom) %>% 
  gather(relationship, wang) %>%
  mutate(relationship = ordered(relationship, levels = relationship_order)) %>% 
  mutate_if(is.numeric, round, 2)

# plot histograms of relatedness 

wgom_plot <- ggplot(sim_wgom, aes(x=wang, fill = relationship, label = relationship, color = relationship)) +
  geom_histogram(binwidth = 0.01, alpha=0.6, position = 'identity', show.legend = F) +
 # geom_vline(aes(xintercept = 0),
  #           color = "black", linetype = "solid", size = 0.75) +
  geom_vline(aes(xintercept = 0.25),
             color = "black", linetype = "solid", size = 0.75) +
  geom_vline(aes(xintercept = 0.5),
             color = "black", linetype = "solid", size = 0.75) +
  geom_vline(aes(xintercept = quantile(sim_hs_wgom, 0.01, na.rm = TRUE)),
             color = "red", linetype = "dotted", size = 0.75) +
  geom_vline(aes(xintercept = quantile(sim_fs_wgom, 0.01, na.rm = TRUE)),
             color = "red", linetype = "dotted", size = 0.75) +
  labs(x = "Relatedness", y = "Number of Pairs") +
  scale_fill_manual(values=col_rel) +
  scale_color_manual(values=col_rel) +
  ggtitle("Western Gulf") +
  theme_standard + 
  theme(plot.title = element_text(hjust = 0.5))

wgom_plot

ggsave(here("related", "results", "wgom_related_sim.png"))

# full sib threshold

fs_thres_wgom <- round(quantile(sim_fs_wgom, 0.01, na.rm = TRUE), 2)
fs_thres_wgom

# half sib threshold

hs_thres_wgom <- round(quantile(sim_hs_wgom, 0.01, na.rm = TRUE), 2)
hs_thres_wgom

```

Identify siblings.

```{r fig.height=5, fig.width=5, message=FALSE, warning=FALSE}

wgom_related <- read.table(here("related", "data", "Demerelate0.7719", "Empirical.relatedness.WGoM.txt")) %>%
  rownames_to_column() %>%
  separate(rowname, into = c("ind1", "ind2"), sep = "_(?=[A-Z])", remove = TRUE) %>%
  mutate_if(is.numeric, round, 5) %>%
  mutate(pop = "wgom")

# match to strata

related1 <- wgom_related %>%
  select(ind1, ind2, wang) %>%
  rename(hiseq_id = ind1) %>%
  left_join(yoy_strata) %>%
  select(c(hiseq_id, sample_id, stage, sex, region, site_shrt, latitude, longitude, year, month, tl, wang))

related2 <- wgom_related %>%
  select(ind2) %>%
  rename(hiseq_id = ind2) %>%
  left_join(yoy_strata) %>%
  select(c(hiseq_id, sample_id, stage, sex, region, site_shrt, latitude, longitude, year, month, tl)) %>%
  rename(hiseq_id_2 = hiseq_id) %>%
  rename(sample_id_2 = sample_id) %>%
  rename(stage_2 = stage) %>%
  rename(sex_2 = sex) %>%
  rename(region_2 = region) %>%
  rename(site_shrt_2 = site_shrt) %>%
  rename(latitude_2 = latitude) %>%
  rename(longitude_2 = longitude) %>%
  rename(year_2 = year) %>%
  rename(month_2 = month) %>%
  rename(tl_2 = tl)

wgom_related <- cbind(related1, related2) %>%
  arrange(desc(wang)) %>% 
  mutate_at("wang", round, 2)

write_csv(wgom_related, here("related", "results", "wgom_related.csv"))

# Plot

ggplot(wgom_related, aes(x = wang)) +
  geom_histogram(binwidth = 0.01, color = "black", fill = "grey") +
  geom_vline(aes(xintercept = mean(wang)),
             color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = 0),
             color = "black", linetype = "solid", size = 1) +
  labs(x = "Relatedness", y = "Number of Pairs") +
  ggtitle("Eastern Gulf") +
  theme_standard + 
  theme(plot.title = element_text(hjust = 0.5))

ggsave(here("related", "results", "wgom_related.png"))

# full siblings

fs_wgom <- wgom_related %>% 
  filter(wang >= fs_thres_wgom)

# half siblings

hs_wgom <- wgom_related %>% 
  filter(wang >= hs_thres_wgom & wang <= fs_thres_wgom)

```

Supporting figure of distribution of simulated values

```{r fig.height=5, fig.width=15, message=FALSE, warning=FALSE}

atl_plot + egom_plot + wgom_plot

ggsave(here("related", "results", "simulations_by_region.png"))

```


### Siblings

Identify randomly and non-randomly sampled siblings.

```{r fig.height=5, fig.width=5, message=FALSE, warning=FALSE}

# collect all siblings

deme_sibs <- rbind(fs_atl, hs_atl, fs_egom, hs_egom, fs_wgom, hs_wgom) %>% 
  mutate(type = case_when(wang > 0.4 ~ "fullsib", type = wang < 0.4 ~ "halfsib")) %>% 
  relocate(type, .after = "wang") %>% 
  arrange(desc(wang))

# non-randomly sampled sibs

nr_kin <- c("TX_201737", "TX_201740", "TX_201741", "TX_201743", "TX_201760", "TX_201735", "TX_47212", "2019-0624-05-02", "JG_2015019")

sibs_non_random <- deme_sibs %>% 
  filter(sample_id %in% nr_kin | sample_id_2 %in% nr_kin) %>%
  arrange(desc(wang))

write_csv(sibs_non_random, here("related", "results", "sibs_non_random.csv"))

# produce list to remove from vcf

sibs_non_random %>%
  select(hiseq_id_2) %>% 
  write_delim(., here("related", "data", "non_random.sib"))

# randomly sampled sibs

sibs_random <- anti_join(deme_sibs, sibs_non_random) %>%
  arrange(desc(wang)) %>% 
  mutate(year_diff = abs(year - year_2)) %>% 
  relocate(year_diff, .after = "type")

write_csv(sibs_random, here("related", "results", "sibs_random.csv"))

```

# Produce Files

Produce files without one individual from non-randomly sampled sibling pairs for population genetic analyses.

```{r}

sibs_non_random <- read_csv(here("related", "results", "sibs_non_random.csv"))

# export 

sibs_non_random %>% 
  select(hiseq_id) %>% 
  distinct() %>% 
  write_delim(here("related", "results", "remove.nonrandom_sibs"))

# produce genepop

gen <- read.genepop(file = here::here("filter", "results", "yoy_final_site.gen"), ncode = 3L, quiet = FALSE)

yoy.gen <- gen.ind.rem.Ind(gen, sibs_non_random$hiseq_id)

yoy_inds <- as_tibble_col(indNames(yoy.gen), column_name = "hiseq_id")

# produce strata

yoy_strata <- read_csv(here("filter", "results", "yoy_final_strata.csv")) %>%
  left_join(yoy_inds, .)

# assign strata

strata(yoy.gen) <- yoy_strata

setPop(yoy.gen) <- ~site_shrt

# export strata and genepop

# remove monomorphic loci

gendiv <- summary(yoy.gen)

dat <- genind2hierfstat(yoy.gen)

gendiv2 <- basic.stats(dat)

# observed heterozygosity per locus

Ho <- as.data.frame(gendiv$Hobs) %>% 
  rownames_to_column("LOCUS") %>%
  rename(Ho = `gendiv$Hobs`)

monomorphic <- filter(Ho, Ho == 0) %>%
  select(LOCUS)

View(monomorphic)

write_delim(monomorphic, here("related", "data", "monomorphic_sibs.loci"), col_names = FALSE, delim = "\t")

remove_loci <- monomorphic$LOCUS

# produce final dataset

yoy.gen <- genind.rem.loci(yoy.gen, remove_loci)

# export genepop by site

setPop(yoy.gen) <- ~site_shrt

yoy.gen$pop <- factor(yoy.gen$pop, levels=site_order_shrt)

writeGenPop(yoy.gen, file.name = here("yoy_site.gen"), comment = "yoy_site.gen")

# Export strata

write_csv(yoy_strata, here("yoy_strata.csv"))

```

Convert genepop file organised by site into `BayeScan` and `Arlequin` files using `pgdspider`. 
  
```{bash, eval=FALSE}

cd ~/Projects/us_blacktips

# bayescan

java -jar /usr/local/bin/PGDSpider2-cli.jar -inputfile yoy_site.gen -inputformat GENEPOP -outputfile fst_outlier/data/yoy_site_bs.txt -outputformat BAYE_SCAN -spid ~/code/genepop_bayescan.spid

# arlequin

java -jar /usr/local/bin/PGDSpider2-cli.jar -inputfile yoy_site.gen -inputformat GENEPOP -outputfile fst_outlier/data/yoy_site.arp -outputformat ARLEQUIN -spid ~/code/genepop_arl.spid

# outflank 

# remove non randomly sampled sibs from filtered vcf

vcftools --vcf filter/data/yoy_filt.recode.vcf --out fst_outlier/data/yoy_temp --remove related/results/remove.nonrandom_sibs --recode --recode-INFO-all

# move loci made monomorphic by removing non-random sibs (if any)

#vcftools --vcf fst_outlier/data/yoy_temp.recode.vcf --not-chr dDocent_Contig_20294 -- out fst_outlier/data/yoy_outflank --recode --recode-INFO-all

vcftools --vcf fst_outlier/data/yoy_temp.recode.vcf --out fst_outlier/data/yoy_outflank --012

```

