---
title: "Population Structure: YOY Blacktips in US. Waters"
author: "DG Swift"
output:
  html_document:
    df_print: paged
    toc: yes
  html_notebook:
    toc: yes
---

Assess population structure using `Arlequin`, PCA, and DAPC.
  
# Environment
  
```{r, message=FALSE}

.libPaths("/usr/lib64/R/library")

# invalidate cache when the package version changes

knitr::opts_chunk$set(
  root.dir = "~/Projects/us_blacktips/pop_str",
	message = FALSE,
	warning = FALSE,
	cache.extra = packageVersion("tint"),
	tidy = FALSE,
	echo = FALSE)

options(htmltools.dir.version = FALSE)

# conflicts

library(conflicted)
conflict_prefer("count", "dplyr")
conflict_prefer("arrange", "dplyr")
conflict_prefer("mutate", "dplyr")
conflict_prefer("select", "dplyr")
conflict_prefer("filter", "dplyr")
conflict_prefer("rename", "dplyr")
conflict_prefer("summarise", "dplyr")
conflict_prefer("summarize", "dplyr")
conflict_prefer("s.label", "adegraphics")
conflict_prefer("s.value", "adegraphics")
conflict_prefer("scalebar", "raster")
conflict_prefer("rename", "dplyr")
conflict_prefer("extract", "raster")
conflict_prefer("degree", "igraph")

# packages

.libPaths("/usr/lib64/R/library")

library(tidyverse)
library(adegenet)
library(ggthemes)
library(tufte)
library(tint)
library(knitr)
library(gdata)
library(zvau)
library(patchwork)
library(here)
library(readxl)

# source scripts

source("~/code/ggplot.R")
source("~/code/genind.R")
source("~/code/PCA.R")
source("~/code/DAPC.R")

# orders, colors, shapes

site_order_shrt <- c("BLB", "SHS", "PRS", "TCB", "WAB", "APB", "MOB", "GAB", "MAB", "SAB", "CCB")

site_order <- c("Bulls_Bay", 
                "St._Helena_Sound",
                "Port_Royal_Sound",
                "Terra_Ceia_Bay",
                "Waccasassa_Bay",
                "Apalachicola_Bay",
                "Mobile_Bay",
                "Galveston_Bay",
                "Matagorda_Bay",
                "San_Antonio_Bay",
                "Corpus_Christi_Bay")

site_order_fig <- c("Bulls Bay", 
                "St. Helena Sound",
                "Port Royal Sound",
                "Terra Ceia Bay",
                "Waccasassa Bay",
                "Apalachicola Bay",
                "Mobile Bay",
                "Galveston Bay",
                "Matagorda Bay",
                "San Antonio Bay",
                "Corpus Christi Bay")

site_order_fig_rev <- c("Corpus Christi Bay",
                        "San Antonio Bay",
                        "Matagorda Bay",
                        "Galveston Bay",
                        "Mobile Bay",
                        "Apalachicola Bay",
                        "Waccasassa Bay",
                        "Terra Ceia Bay",
                        "Port Royal Sound",
                        "St. Helena Sound",
                        "Bulls Bay")

site_order_fig_map <- c("Galveston Bay",
                        "Matagorda Bay",
                        "San Antonio Bay",
                        "Corpus Christi Bay",
                        "Mobile Bay",
                        "Apalachicola Bay",
                        "Waccasassa Bay",
                        "Terra Ceia Bay",
                        "Bulls Bay",
                        "St. Helena Sound",
                        "Port Royal Sound")

site_order_shrt_rev <- c("CCB", "SAB", "MAB", "GAB", "MOB", "APB", "WAB", "TCB", "PRS", "SHS", "BLB")

region_order <- c("Atl", "EGoM", "WGoM")

region_order_fig <- c("Atlantic", "Eastern Gulf", "Western Gulf")

region_order_fig_rev <- c("Western Gulf", "Eastern Gulf", "Atlantic")

site_col <- c("#7f0000", "#b30000", "#d7301f", "#4d004b", "#810f7c", "#88419d", "#8c96c6", "#9ecae1", "#4292c6", "#08519c", "#08306b")

site_col_rev <- c("#08306b", "#08519c", "#4292c6", "#9ecae1", "#8c96c6", "#88419d", "#810f7c", "#4d004b", "#d7301f", "#b30000", "#7f0000")

site_col_map <- c("#9ecae1", "#4292c6", "#08519c", "#08306b", "#8c96c6", "#88419d", "#810f7c", "#4d004b", "#d7301f", "#b30000", "#7f0000")
              
shape11 <- c(21, 21, 21, 23, 23, 23, 23, 24, 24, 24, 24)

shape11_rev <- c(24, 24, 24, 24, 23, 23, 23, 23, 21, 21, 21)

shape3 <- c(21, 23, 24)

shape3_rev <- c(24, 23, 21)

region_col <- c('#b30000','#810f7c','#08519c')

lib_col <- c("#A6CEE3", "#1F78B4", "#B2DF8A", "#33A02C", "#FB9A99", "#E31A1C", "#FDBF6F", "#FF7F00", "#CAB2D6", "#6A3D9A", "#FFFF33")

options(scipen=10000)

```

# Arlequin

Execute AMOVA and post-hoc pairwise FST tests among regions and sites.

Rename `pop` in `.arp` files with site name. Prep `.arp`files for locus-by-locus pairwise FST tests and insert structure formats. 

```{bash, eval = FALSE}

cd /home/dswift/Projects/us_blacktips/pop_str/data

# rename sites

ls yoy_*.arp | while read i; do sed -i 's/"pop_1"/"BLB"/g' $i $i; sed -i 's/"pop_2"/"SHS"/g' $i $i; sed -i 's/"pop_3"/"PRS"/g' $i $i; sed -i 's/"pop_4"/"TCB"/g' $i $i; sed -i 's/"pop_5"/"WAB"/g' $i $i; sed -i 's/"pop_6"/"APB"/g' $i $i; sed -i 's/"pop_7"/"MOB"/g' $i $i; sed -i 's/"pop_8"/"GAB"/g' $i $i; sed -i 's/"pop_9"/"MAB"/g' $i $i; sed -i 's/"pop_10"/"SAB"/g' $i $i; sed -i 's/"pop_11"/"CCB"/g' $i $i; done

# amova and pwfst

cd /home/dswift/Projects/us_blacktips/pop_str/data/amova

## produce .arp files with different structures

rm arl_run.txt randseed.txt

ls *.txt | cut -f1 -d"." | while read i; do cat ../yoy_neu.arp $i.txt > $i"_neu".arp ; done
ls *.txt | cut -f1 -d"." | while read i; do cat ../yoy_ada.arp $i.txt > $i"_ada".arp ; done
ls *.txt | cut -f1 -d"." | while read i; do cat ../yoy_wosibs_neu.arp $i.txt > $i"_wosibs_neu".arp ; done
ls *.txt | cut -f1 -d"." | while read i; do cat ../yoy_wosibs_ada.arp $i.txt > $i"_wosibs_ada".arp ; done

## execute

ls *.arp | while read i; do arlecore3522_64bit $i amova_pwfst.ars ; done

## move output files

mv */*.xml ../../results/amova/.

# lbl pwfst by site

cd /home/dswift/Projects/us_blacktips/pop_str/data/lbl_pwfst

## produce .arp files with different structures

rm arl_run.txt randseed.txt

ls *.txt | cut -f1 -d"." | while read i; do cat ../yoy_neu.arp $i.txt > $i"_neu".arp ; done
ls *.txt | cut -f1 -d"." | while read i; do cat ../yoy_ada.arp $i.txt > $i"_ada".arp ; done
ls *.txt | cut -f1 -d"." | while read i; do cat ../yoy_wosibs_neu.arp $i.txt > $i"_wosibs_neu".arp ; done
ls *.txt | cut -f1 -d"." | while read i; do cat ../yoy_wosibs_ada.arp $i.txt > $i"_wosibs_ada".arp ; done

## execute

ls *.arp | while read i; do arlecore3522_64bit $i lbl_pwfst_no_missing.ars ; done

## move output files

mv */*.xml ../../results/lbl_pwfst/.

cd /home/dswift/Projects/us_blacktips/pop_str/results/lbl_pwfst

grep -A 2 'Average\|Va and FST' *str_neu.xml > lbl_pwfst_neu_results.txt

grep -A 2 'Average\|Va and FST' *str_ada.xml > lbl_pwfst_ada_results.txt

grep -A 2 'Average\|Va and FST' *str_wosibs_neu.xml > lbl_pwfst_wosibs_neu_results.txt

grep -A 2 'Average\|Va and FST' *str_wosibs_ada.xml > lbl_pwfst_wosibs_ada_results.txt

```

## Results

### Pairwise FST and p-values

Round up negative pwFST values and correct p-values for multiple testing using Benjamini-Hochberg.

```{r, warning = FALSE, message = FALSE}

yoy_strata <- read_csv(here("yoy_strata.csv"))

# read in csv files

lbl_pwfst_mtdna <- read_xlsx(here("pop_str", "results", "lbl_pwfst", "lbl_pwfst_results.xlsx"), sheet = "mtdna_wosibs") %>%
  column_to_rownames("X")

lbl_pwfst_neu <- read_xlsx(here("pop_str", "results", "lbl_pwfst", "lbl_pwfst_results.xlsx"), sheet = "neu_wosibs") %>%
  column_to_rownames("X")

lbl_pwfst_ada <- read_xlsx(here("pop_str", "results", "lbl_pwfst", "lbl_pwfst_results.xlsx"), sheet = "ada_wosibs") %>%
  column_to_rownames("X")

# extract pwfst values as vectors

lbl_mtdna_fst <- as.numeric(lowerTriangle(lbl_pwfst_mtdna, diag = TRUE, byrow = FALSE))

lbl_neu_fst <- as.numeric(lowerTriangle(lbl_pwfst_neu, diag = TRUE, byrow = FALSE))

lbl_ada_fst <- as.numeric(lowerTriangle(lbl_pwfst_ada, diag = TRUE, byrow = FALSE))

# produce df of pwfst values, replace negative values with 0, and round to 5 sf

fst_df <- as.data.frame(cbind(lbl_mtdna_fst, lbl_neu_fst, lbl_ada_fst)) %>%
  mutate_if(is.numeric, funs(ifelse(. < 0, 0, .))) %>%
  mutate_all(round, 5)

# extract p-values as vectors

lbl_mtdna_pval <- as.numeric(upperTriangle(lbl_pwfst_mtdna, diag = TRUE, byrow = TRUE))

lbl_neu_pval <- as.numeric(upperTriangle(lbl_pwfst_neu, diag = TRUE, byrow = TRUE))

lbl_ada_pval <- as.numeric(upperTriangle(lbl_pwfst_ada, diag = TRUE, byrow = TRUE))

# collect all p values in a df and correct for multiple tests with BH

pval_df <- as.data.frame(cbind(lbl_mtdna_pval, lbl_neu_pval, lbl_ada_pval)) %>%
  mutate_all(funs(fdr = p.adjust(., method = "BH"))) %>%
  mutate(across(1:6, round, 5))

```

Produce df of pwFST, corrected, and uncorrected p-values for each dataset.

```{r}

# produce vector of site-site comparisons

comp <- unmatrix(lbl_pwfst_neu, byrow = TRUE) %>%
  as.data.frame(.) %>%
  #filter(. != "-") %>%
  rownames_to_column("comp") %>%
  select(comp) %>%
  separate(comp, into = c("site1", "site2"), sep = "[:]", remove = FALSE) %>%
  #slice(1:8, 10:16, 19:24, 28:32, 37:40, 46:48, 55:56, 64)
  slice(1:11, 13:22, 25:33, 37:44, 49:55, 61:66, 73:77, 85:88, 97:99, 109:110, 121)

# import shark distance between sites to combine with fst and p values

site_shark_dist <- read_csv(here("rda", "data", "site_shark_dist.csv"))

# lbl mtdna

lbl_mtdna <- as.data.frame(cbind(comp, fst_df$lbl_mtdna_fst, pval_df$lbl_mtdna_pval, pval_df$lbl_mtdna_pval_fdr)) %>%
  rename(pwFST = "fst_df$lbl_mtdna_fst") %>%
  rename(pval = "pval_df$lbl_mtdna_pval") %>%
  rename(cor_pval = "pval_df$lbl_mtdna_pval_fdr") %>%
  mutate(site1 = ordered(site1, levels = site_order_shrt)) %>%
  mutate(site2 = ordered(site2, levels = site_order_shrt_rev)) %>%
  mutate(region1 = case_when(site1 %in% c("BLB", "SHS", "PRS") ~ "Atl",
                             site1 %in% c("TCB", "WAB", "APB", "MOB") ~ "EGoM",
                             site1 %in% c("GAB", "MAB", "SAB", "CCB") ~ "WGoM")) %>%
  mutate(region2 = case_when(site2 %in% c("BLB", "SHS", "PRS") ~ "Atl",
                             site2 %in% c("TCB", "WAB", "APB", "MOB") ~ "EGoM",
                             site2 %in% c("GAB", "MAB", "SAB", "CCB") ~ "WGoM")) %>%
  mutate(Comparison = case_when(region1 == "Atl" & region2 == "EGoM" ~ "Atl-EGoM", 
                                region1 == "Atl" & region2 == "WGoM" ~ "Atl-WGoM", 
                                region1 == "EGoM" & region2 =="WGoM" ~ "EGoM-WGoM",
                                region1 == "Atl" & region2 == "Atl" ~ "Atl-Atl", 
                                region1 == "EGoM" & region2 == "EGoM" ~ "EGoM-EGoM", 
                                region1 == "WGoM" & region2 =="WGoM" ~ "WGoM-WGoM")) %>%
  select(-c(region1, region2)) %>%
  mutate(sign = case_when(cor_pval > 0.05 ~ "",
                          cor_pval < 0.05 & cor_pval >= 0.01 ~ "*",
                          cor_pval < 0.01 & cor_pval >= 0.001 ~ "**",
                          cor_pval <= 0.001 ~ "***")) %>%
  cbind(site_shark_dist) %>%
  mutate(dist = site_shark_dist/1000) %>% 
  select(-site_shark_dist) %>% 
  drop_na(pwFST, Comparison)

write_csv(lbl_mtdna, here("pop_str", "results", "lbl_pwfst", "lbl_mtdna_df.csv"))

# lbl neu

lbl_neu <- as.data.frame(cbind(comp, fst_df$lbl_neu_fst, pval_df$lbl_neu_pval, pval_df$lbl_neu_pval_fdr)) %>%
  rename(pwFST = "fst_df$lbl_neu_fst") %>%
  rename(pval = "pval_df$lbl_neu_pval") %>%
  rename(cor_pval = "pval_df$lbl_neu_pval_fdr") %>%
  mutate(site1 = ordered(site1, levels = site_order_shrt)) %>%
  mutate(site2 = ordered(site2, levels = site_order_shrt_rev)) %>%
  mutate(region1 = case_when(site1 %in% c("BLB", "SHS", "PRS") ~ "Atl",
                             site1 %in% c("TCB", "WAB", "APB", "MOB") ~ "EGoM",
                             site1 %in% c("GAB", "MAB", "SAB", "CCB") ~ "WGoM")) %>%
  mutate(region2 = case_when(site2 %in% c("BLB", "SHS", "PRS") ~ "Atl",
                             site2 %in% c("TCB", "WAB", "APB", "MOB") ~ "EGoM",
                             site2 %in% c("GAB", "MAB", "SAB", "CCB") ~ "WGoM")) %>%
  mutate(Comparison = case_when(region1 == "Atl" & region2 == "EGoM" ~ "Atl-EGoM", 
                                region1 == "Atl" & region2 == "WGoM" ~ "Atl-WGoM", 
                                region1 == "EGoM" & region2 =="WGoM" ~ "EGoM-WGoM",
                                region1 == "Atl" & region2 == "Atl" ~ "Atl-Atl", 
                                region1 == "EGoM" & region2 == "EGoM" ~ "EGoM-EGoM", 
                                region1 == "WGoM" & region2 =="WGoM" ~ "WGoM-WGoM")) %>%
  select(-c(region1, region2)) %>%
  mutate(sign = case_when(cor_pval > 0.05 ~ "",
                          cor_pval < 0.05 & cor_pval >= 0.01 ~ "*",
                          cor_pval < 0.01 & cor_pval >= 0.001 ~ "**",
                          cor_pval <= 0.001 ~ "***")) %>%
  cbind(site_shark_dist) %>%
  mutate(dist = site_shark_dist/1000) %>% 
  select(-site_shark_dist) %>% 
  drop_na(pwFST, Comparison)

write_csv(lbl_neu, here("pop_str", "results", "lbl_pwfst", "lbl_neu_df.csv"))

# lbl ada

lbl_ada <- as.data.frame(cbind(comp, fst_df$lbl_ada_fst, pval_df$lbl_ada_pval, pval_df$lbl_ada_pval_fdr)) %>%
  rename(pwFST = "fst_df$lbl_ada_fst") %>%
  rename(pval = "pval_df$lbl_ada_pval") %>%
  rename(cor_pval = "pval_df$lbl_ada_pval_fdr") %>%
  mutate(site1 = ordered(site1, levels = site_order_shrt)) %>%
  mutate(site2 = ordered(site2, levels = site_order_shrt_rev)) %>%
  mutate(region1 = case_when(site1 %in% c("BLB", "SHS", "PRS") ~ "Atl",
                             site1 %in% c("TCB", "WAB", "APB", "MOB") ~ "EGoM",
                             site1 %in% c("GAB", "MAB", "SAB", "CCB") ~ "WGoM")) %>%
  mutate(region2 = case_when(site2 %in% c("BLB", "SHS", "PRS") ~ "Atl",
                             site2 %in% c("TCB", "WAB", "APB", "MOB") ~ "EGoM",
                             site2 %in% c("GAB", "MAB", "SAB", "CCB") ~ "WGoM")) %>%
  mutate(Comparison = case_when(region1 == "Atl" & region2 == "EGoM" ~ "Atl-EGoM", 
                                region1 == "Atl" & region2 == "WGoM" ~ "Atl-WGoM", 
                                region1 == "EGoM" & region2 =="WGoM" ~ "EGoM-WGoM",
                                region1 == "Atl" & region2 == "Atl" ~ "Atl-Atl", 
                                region1 == "EGoM" & region2 == "EGoM" ~ "EGoM-EGoM", 
                                region1 == "WGoM" & region2 =="WGoM" ~ "WGoM-WGoM")) %>%
  select(-c(region1, region2)) %>%
  mutate(sign = case_when(cor_pval > 0.05 ~ "",
                          cor_pval < 0.05 & cor_pval >= 0.01 ~ "*",
                          cor_pval < 0.01 & cor_pval >= 0.001 ~ "**",
                          cor_pval <= 0.001 ~ "***")) %>%
  cbind(site_shark_dist) %>%
  mutate(dist = site_shark_dist/1000) %>% 
  select(-site_shark_dist) 

write_csv(lbl_ada, here("pop_str", "results", "lbl_pwfst", "lbl_ada_df.csv"))

# export site comparisons from pairwise fst tests

site_comp <- select(lbl_ada, comp)

#write.csv(site_comp, "../Results/site_comparison.csv", row.names = FALSE)

```

#### mtDNA ΦST Vs. Geographic Distance

Assess the relationship between pairwise FST and geographic distance between sites between and within regions.

```{r message=FALSE, warning=FALSE}

# linear regression by group

## Atlantic-Eastern Gulf

temp <- lbl_mtdna %>%
  filter(Comparison == "Atl-EGoM") %>% 
  drop_na(pwFST)

lm_mtdna_atl_egom <- lm(temp$dist ~ temp$pwFST)

summary(lm_mtdna_atl_egom)

## Atlantic-Western Gulf

temp <- lbl_mtdna %>%
  filter(Comparison == "Atl-WGoM") %>% 
  drop_na(pwFST)

lm_mtdna_atl_wgom <- lm(temp$dist ~ temp$pwFST) 

summary(lm_mtdna_atl_wgom)

## Within Gulf

temp <- lbl_mtdna %>%
  filter(Comparison == "EGoM-WGoM") %>% 
  drop_na(pwFST)

lm_mtdna_egom_wgom <- lm(temp$dist ~ temp$pwFST)

summary(lm_mtdna_egom_wgom)

```

Plot regional comparisons

```{r fig.height=5, fig.width=15, message=FALSE, warning=FALSE}

# plot

temp <- lbl_mtdna %>% 
  filter(Comparison == "Atl-EGoM" | Comparison == "Atl-WGoM" | Comparison == "EGoM-WGoM")

dat_text <- data.frame(
  label = c("ns", "ns", "*"),
  Comparison = c("Atl-EGoM", "Atl-WGoM", "EGoM-WGoM"),
  x = c(150, 150, 150),
  y = c(0.275, 0.275, 0.275))

ggplot(data = temp, aes(x = dist, y = pwFST), show.legend = F) +
  geom_point(size = 2, alpha = 0.75) +
  geom_smooth(method = lm, size = 1, alpha = 0.5) +
#  scale_x_continuous(limits = c(0, 2650), breaks = seq(0, 2500, 500), labels= seq(0, 2500, 500)) +
 # scale_y_continuous(limits = c(-0.0005, 0.3), breaks = seq(0, 0.3, 0.1), labels= seq(0, 0.3, 0.1), ) +
#  facet_wrap(~ Group, ncol = 3) +
  facet_grid(. ~ Comparison, scales = "fixed", space = "fixed") +
  labs(x = "Geographic Distance (km)", y = expression(paste("Pairwise ", Φ[ST]))) +
  #ggtitle("A") +
  guides(fill = F, shape = F, color = F) +
  theme_standard +
  theme(legend.text=element_text(size=12), legend.title=element_text(size=12), legend.position = "right", legend.box = "vertical", plot.title = element_text(size = 26), axis.text=element_text(size=12)) +
  geom_text(
  data    = dat_text,
  size = c(5, 5, 10),
  color = "red",
  mapping = aes(x = x, y = y, label = label))

```

Gulf only plots

```{r fig.height=5, fig.width=5, message=FALSE, warning=FALSE}

# egom-wgom

temp <- lbl_mtdna %>% 
  select(-Comparison) %>% 
  mutate(region1 = case_when(site1 %in% c("BLB", "SHS", "PRS") ~ "Atl",
                             site1 %in% c("TCB", "WAB", "APB", "MOB") ~ "EGoM",
                             site1 %in% c("GAB", "MAB", "SAB", "CCB") ~ "WGoM")) %>%
  mutate(region2 = case_when(site2 %in% c("BLB", "SHS", "PRS") ~ "Atl",
                             site2 %in% c("TCB", "WAB", "APB", "MOB") ~ "EGoM",
                             site2 %in% c("GAB", "MAB", "SAB", "CCB") ~ "WGoM")) %>%
  mutate(Comparison = case_when(region1 == "EGoM" & region2 =="WGoM" ~ "EGoM-WGoM")) %>% 
  select(-c(region1, region2)) %>%
  drop_na(pwFST, Comparison)

dat_text <- data.frame(
  label = "*",
  x = 1000,
  y = 0.225)

ggplot(data = temp, aes(x = dist, y = pwFST), show.legend = F) +
  geom_point(size = 2, alpha = 0.75) +
  geom_smooth(method = lm, size = 1, alpha = 0.5) +
  #scale_x_continuous(limits = c(500, 1500), breaks = seq(500, 1500, 250), labels= seq(500, 1500, 250)) +
  #scale_y_continuous(limits = c(-0.05, 0.25), breaks = seq(0, 0.25, 0.05), labels= seq(0, 0.25, 0.05)) +
  labs(x = "Geographic Distance (km)", y = expression(paste("Pairwise ", Φ[ST]))) +
  #ggtitle("A") +
  guides(fill = F, shape = F, color = F) +
  theme_standard +
  theme(legend.text=element_text(size=12), legend.title=element_text(size=12), legend.position = "right", legend.box = "vertical", plot.title = element_text(size = 20), axis.text=element_text(size=12)) +
  geom_text(
  data  = dat_text,
  size = 8,
  color = "red",
  mapping = aes(x = x, y = y, label = label))

# within gulf

temp <- lbl_mtdna %>% 
  select(-Comparison) %>% 
  mutate(region1 = case_when(site1 %in% c("BLB", "SHS", "PRS") ~ "Atl",
                             site1 %in% c("TCB", "WAB", "APB", "MOB") ~ "EGoM",
                             site1 %in% c("GAB", "MAB", "SAB", "CCB") ~ "WGoM")) %>%
  mutate(region2 = case_when(site2 %in% c("BLB", "SHS", "PRS") ~ "Atl",
                             site2 %in% c("TCB", "WAB", "APB", "MOB") ~ "EGoM",
                             site2 %in% c("GAB", "MAB", "SAB", "CCB") ~ "WGoM")) %>%
  mutate(Comparison = case_when(region1 == "EGoM" & region2 =="WGoM" | region1 == "EGoM" & region2 =="EGoM" | region1 == "WGoM" & region2 =="WGoM" ~ "Within Gulf")) %>% 
  select(-c(region1, region2)) %>%
  drop_na(pwFST, Comparison)

dat_text <- data.frame(
  label = "*",
  x = 1000,
  y = 0.225)

ggplot(data = temp, aes(x = dist, y = pwFST), show.legend = F) +
  geom_point(size = 2, alpha = 0.75) +
  geom_smooth(method = lm, size = 1, alpha = 0.5) +
  #scale_x_continuous(limits = c(500, 1500), breaks = seq(500, 1500, 250), labels= seq(500, 1500, 250)) +
  #scale_y_continuous(limits = c(-0.05, 0.25), breaks = seq(0, 0.25, 0.05), labels= seq(0, 0.25, 0.05)) +
  labs(x = "Geographic Distance (km)", y = expression(paste("Pairwise ", Φ[ST]))) +
  #ggtitle("B") +
  guides(fill = F, shape = F, color = F) +
  theme_standard +
  theme(legend.text=element_text(size=12), legend.title=element_text(size=12), legend.position = "right", legend.box = "vertical", plot.title = element_text(size = 20), axis.text=element_text(size=12)) +
  geom_text(
  data  = dat_text,
  size = 8,
  color = "red",
  mapping = aes(x = x, y = y, label = label))

```

#### Neutral SNPs FST Vs. Geographic Distance

Assess the relationship between pairwise FST and geographic distance between sites between and within regions.

```{r message=FALSE, warning=FALSE}

# linear regression by group

## Atlantic-Eastern Gulf

temp <- lbl_neu %>%
  filter(Comparison == "Atl-EGoM") %>% 
  drop_na(pwFST)

lm_neu_atl_egom <- lm(temp$dist ~ temp$pwFST)

summary(lm_neu_atl_egom)

## Atlantic-Western Gulf

temp <- lbl_neu %>%
  filter(Comparison == "Atl-WGoM") %>% 
  drop_na(pwFST)

lm_neu_atl_wgom <- lm(temp$dist ~ temp$pwFST) 

summary(lm_neu_atl_wgom)

## Within Gulf

temp <- lbl_neu %>%
  filter(Comparison == "EGoM-WGoM") %>% 
  drop_na(pwFST)

lm_neu_egom_wgom <- lm(temp$dist ~ temp$pwFST)

summary(lm_neu_egom_wgom)

```

Plot regional comparisons

```{r fig.height=5, fig.width=15, message=FALSE, warning=FALSE}

# plot

temp <- lbl_neu %>% 
  filter(Comparison == "Atl-EGoM" | Comparison == "Atl-WGoM" | Comparison == "EGoM-WGoM")

dat_text <- data.frame(
  label = c("ns", "ns", "**"),
  Comparison = c("Atl-EGoM", "Atl-WGoM", "EGoM-WGoM"),
  x = c(400, 400, 400),
  y = c(0.004, 0.004, 0.004))

ggplot(data = temp, aes(x = dist, y = pwFST), show.legend = F) +
  geom_point(size = 2, alpha = 0.75) +
  geom_smooth(method = lm, size = 1, alpha = 0.5) +
 # scale_x_continuous(limits = c(0, 2650), breaks = seq(0, 2500, 500), labels= seq(0, 2500, 500)) +
  #scale_y_continuous(limits = c(-0.0005, 0.3), breaks = seq(0, 0.3, 0.1), labels= seq(0, 0.3, 0.1), ) +
#  facet_wrap(~ Group, ncol = 3) +
  facet_grid(. ~ Comparison, scales = "fixed", space = "fixed") +
  labs(x = "Geographic Distance (km)", y = expression(paste("Pairwise Neutral ", italic(F)[ST]))) +
  ggtitle("A") +
  guides(fill = F, shape = F, color = F) +
  theme_standard +
  theme(legend.text=element_text(size=12), legend.title=element_text(size=12), legend.position = "right", legend.box = "vertical", plot.title = element_text(size = 26), axis.text=element_text(size=12)) +
  geom_text(
  data    = dat_text,
  size = c(5, 5, 10),
  color = "red",
  mapping = aes(x = x, y = y, label = label))

```

Gulf only plots

```{r fig.height=5, fig.width=5, message=FALSE, warning=FALSE}

# egom-wgom

temp <- lbl_neu %>% 
  select(-Comparison) %>% 
  mutate(region1 = case_when(site1 %in% c("BLB", "SHS", "PRS") ~ "Atl",
                             site1 %in% c("TCB", "WAB", "APB", "MOB") ~ "EGoM",
                             site1 %in% c("GAB", "MAB", "SAB", "CCB") ~ "WGoM")) %>%
  mutate(region2 = case_when(site2 %in% c("BLB", "SHS", "PRS") ~ "Atl",
                             site2 %in% c("TCB", "WAB", "APB", "MOB") ~ "EGoM",
                             site2 %in% c("GAB", "MAB", "SAB", "CCB") ~ "WGoM")) %>%
  mutate(Comparison = case_when(region1 == "EGoM" & region2 =="WGoM" ~ "EGoM-WGoM")) %>% 
  select(-c(region1, region2)) %>%
  drop_na(pwFST, Comparison)

dat_text <- data.frame(
  label = "*",
  x = 1000,
  y = 0.001)

ggplot(data = temp, aes(x = dist, y = pwFST), show.legend = F) +
  geom_point(size = 2, alpha = 0.75) +
  geom_smooth(method = lm, size = 1, alpha = 0.5) +
  scale_x_continuous(limits = c(700, 1500), breaks = seq(750, 1500, 250), labels= seq(750, 1500, 250)) +
  #scale_y_continuous(limits = c(-0.05, 0.25), breaks = seq(0, 0.25, 0.05), labels= seq(0, 0.25, 0.05)) +
  labs(x = "Geographic Distance (km)", y = expression(paste("Pairwise Neutral ", italic(F)[ST]))) +
  ggtitle("D") +
  guides(fill = F, shape = F, color = F) +
  theme_standard +
  theme(legend.text=element_text(size=12), legend.title=element_text(size=12), legend.position = "right", legend.box = "vertical", plot.title = element_text(size = 20), axis.text=element_text(size=12)) +
  geom_text(
  data  = dat_text,
  size = 8,
  color = "red",
  mapping = aes(x = x, y = y, label = label))

# within gulf

temp <- lbl_neu %>% 
  select(-Comparison) %>% 
  mutate(region1 = case_when(site1 %in% c("BLB", "SHS", "PRS") ~ "Atl",
                             site1 %in% c("TCB", "WAB", "APB", "MOB") ~ "EGoM",
                             site1 %in% c("GAB", "MAB", "SAB", "CCB") ~ "WGoM")) %>%
  mutate(region2 = case_when(site2 %in% c("BLB", "SHS", "PRS") ~ "Atl",
                             site2 %in% c("TCB", "WAB", "APB", "MOB") ~ "EGoM",
                             site2 %in% c("GAB", "MAB", "SAB", "CCB") ~ "WGoM")) %>%
  mutate(Comparison = case_when(region1 == "EGoM" & region2 =="WGoM" | region1 == "EGoM" & region2 =="EGoM" | region1 == "WGoM" & region2 =="WGoM" ~ "Within Gulf")) %>% 
  select(-c(region1, region2)) %>%
  drop_na(pwFST, Comparison)

dat_text <- data.frame(
  label = "*",
  x = 1000,
  y = 0.001)

ggplot(data = temp, aes(x = dist, y = pwFST), show.legend = F) +
  geom_point(size = 2, alpha = 0.75) +
  geom_smooth(method = lm, size = 1, alpha = 0.5) +
  #scale_x_continuous(limits = c(500, 1500), breaks = seq(500, 1500, 250), labels= seq(500, 1500, 250)) +
  #scale_y_continuous(limits = c(-0.05, 0.25), breaks = seq(0, 0.25, 0.05), labels= seq(0, 0.25, 0.05)) +
  labs(x = "Geographic Distance (km)", y = expression(paste("Pairwise Neutral ", italic(F)[ST]))) +
 # ggtitle("B") +
  guides(fill = F, shape = F, color = F) +
  theme_standard +
  theme(legend.text=element_text(size=12), legend.title=element_text(size=12), legend.position = "right", legend.box = "vertical", plot.title = element_text(size = 20), axis.text=element_text(size=12)) +
  geom_text(
  data  = dat_text,
  size = 8,
  color = "red",
  mapping = aes(x = x, y = y, label = label))

```

#### Adaptive SNPs FST Vs. Latitude

Assess the relationship between pairwise FST and latitude between sites.

```{r message=FALSE, warning=FALSE}

lat_diff <- read_csv(here("rda", "data", "lat_diff.csv"))

fst_ada_df <- lbl_ada %>% 
  cbind(lat_diff) %>% 
  drop_na(pwFST, Comparison)

# linear regression by group

## Atlantic-Eastern Gulf

temp <- fst_ada_df %>%
  filter(Comparison == "Atl-EGoM") %>% 
  drop_na(pwFST)

lm_ada_atl_egom <- lm(temp$lat_diff ~ temp$pwFST)

summary(lm_ada_atl_egom)

## Atlantic-Western Gulf

temp <- fst_ada_df %>%
  filter(Comparison == "Atl-WGoM") %>% 
  drop_na(pwFST)

lm_ada_atl_wgom <- lm(temp$lat_diff ~ temp$pwFST) 

summary(lm_ada_atl_wgom)

## Within Gulf

temp <- fst_ada_df %>%
  filter(Comparison == "EGoM-WGoM") %>% 
  drop_na(pwFST)

lm_ada_egom_wgom <- lm(temp$lat_diff ~ temp$pwFST)

summary(lm_ada_egom_wgom)

```

Plot regional comparisons

```{r fig.height=5, fig.width=15, message=FALSE, warning=FALSE}

# plot

temp <- fst_ada_df %>% 
  filter(Comparison == "Atl-EGoM" | Comparison == "Atl-WGoM" | Comparison == "EGoM-WGoM")

dat_text <- data.frame(
  label = c("ns", "ns", "**"),
  Comparison = c("Atl-EGoM", "Atl-WGoM", "EGoM-WGoM"),
  x = c(400, 400, 400),
  y = c(0.004, 0.004, 0.004))

ggplot(data = temp, aes(x = lat_diff, y = pwFST), show.legend = F) +
  geom_point(size = 2, alpha = 0.75) +
  geom_smooth(method = lm, size = 1, alpha = 0.5) +
  #scale_x_continuous(limits = c(0, 3), breaks = seq(0, 3, 1), labels= seq(0, 3, 1)) +
  #scale_y_continuous(limits = c(-0.0005, 0.3), breaks = seq(0, 0.3, 0.1), labels= seq(0, 0.3, 0.1), ) +
#  facet_wrap(~ Group, ncol = 3) +
  facet_grid(. ~ Comparison, scales = "fixed", space = "fixed") +
  labs(x = "Absolute Difference in Degrees Latitude", y = expression(paste("Pairwise Adaptive ", italic(F)[ST]))) +
  ggtitle("A") +
  guides(fill = F, shape = F, color = F) +
  theme_standard +
  theme(legend.text=element_text(size=12), legend.title=element_text(size=12), legend.position = "right", legend.box = "vertical", plot.title = element_text(size = 26), axis.text=element_text(size=12))

```

egom-wgom only plots

```{r fig.height=5, fig.width=5, message=FALSE, warning=FALSE}

# egom-wgom

temp <- fst_ada_df %>% 
  select(-Comparison) %>% 
  mutate(region1 = case_when(site1 %in% c("BLB", "SHS", "PRS") ~ "Atl",
                             site1 %in% c("TCB", "WAB", "APB", "MOB") ~ "EGoM",
                             site1 %in% c("GAB", "MAB", "SAB", "CCB") ~ "WGoM")) %>%
  mutate(region2 = case_when(site2 %in% c("BLB", "SHS", "PRS") ~ "Atl",
                             site2 %in% c("TCB", "WAB", "APB", "MOB") ~ "EGoM",
                             site2 %in% c("GAB", "MAB", "SAB", "CCB") ~ "WGoM")) %>%
  mutate(Comparison = case_when(region1 == "EGoM" & region2 =="WGoM" ~ "EGoM-WGoM")) %>% 
  select(-c(region1, region2)) %>%
  drop_na(pwFST, Comparison)

dat_text <- data.frame(
  label = "*",
  x = 1000,
  y = 0.001)

ggplot(data = temp, aes(x = lat_diff, y = pwFST), show.legend = F) +
  geom_point(size = 2, alpha = 0.75) +
  geom_smooth(method = lm, size = 1, alpha = 0.5) +
  #scale_x_continuous(limits = c(500, 1500), breaks = seq(500, 1500, 250), labels= seq(500, 1500, 250)) +
  scale_y_continuous(limits = c(-0.0025, 0.025), breaks = seq(0, 0.025, 0.005), labels= seq(0, 0.025, 0.005)) +
  labs(x = "Absolute Difference in Degrees Latitude", y = expression(paste("Pairwise Adaptive ", italic(F)[ST]))) +
  ggtitle("H") +
  guides(fill = F, shape = F, color = F) +
  theme_standard +
  theme(legend.text=element_text(size=12), legend.title=element_text(size=12), legend.position = "right", legend.box = "vertical", plot.title = element_text(size = 20), axis.text=element_text(size=12))

```

Gulf only plots

```{r fig.height=5, fig.width=5, message=FALSE, warning=FALSE}

# within gulf

temp <- fst_ada_df %>% 
  select(-Comparison) %>% 
  mutate(region1 = case_when(site1 %in% c("BLB", "SHS", "PRS") ~ "Atl",
                             site1 %in% c("TCB", "WAB", "APB", "MOB") ~ "EGoM",
                             site1 %in% c("GAB", "MAB", "SAB", "CCB") ~ "WGoM")) %>%
  mutate(region2 = case_when(site2 %in% c("BLB", "SHS", "PRS") ~ "Atl",
                             site2 %in% c("TCB", "WAB", "APB", "MOB") ~ "EGoM",
                             site2 %in% c("GAB", "MAB", "SAB", "CCB") ~ "WGoM")) %>%
  mutate(Comparison = case_when(region1 == "EGoM" & region2 =="WGoM" | region1 == "EGoM" & region2 =="EGoM" | region1 == "WGoM" & region2 =="WGoM" ~ "Within Gulf")) %>% 
  select(-c(region1, region2)) %>%
  drop_na(pwFST, Comparison)

dat_text <- data.frame(
  label = "*",
  x = 1000,
  y = 0.001)



ggplot(data = temp, aes(x = lat_diff, y = pwFST), show.legend = F) +
  geom_point(size = 2, alpha = 0.75) +
  geom_smooth(method = lm, size = 1, alpha = 0.5) +
  #scale_x_continuous(limits = c(500, 1500), breaks = seq(500, 1500, 250), labels= seq(500, 1500, 250)) +
  #scale_y_continuous(limits = c(-0.05, 0.25), breaks = seq(0, 0.25, 0.05), labels= seq(0, 0.25, 0.05)) +
  labs(x = "Absolute Difference in Degrees Latitude", y = expression(paste("Pairwise Adaptive ", italic(F)[ST]))) +
  #ggtitle("B") +
  guides(fill = F, shape = F, color = F) +
  theme_standard +
  theme(legend.text=element_text(size=12), legend.title=element_text(size=12), legend.position = "right", legend.box = "vertical", plot.title = element_text(size = 20), axis.text=element_text(size=12))

# linear regression by group

## Within Gulf

temp2 <- temp %>%
  filter(Comparison == "Within Gulf") %>% 
  drop_na(pwFST)

lm_ada_within_gulf <- lm(temp2$lat_diff ~ temp2$pwFST)

summary(lm_ada_within_gulf)


```

#### Adaptive SNPs FST Vs. Longitude

```{r fig.height=3.5, fig.width=9.5, message=FALSE, warning=FALSE}

# import differences in longitude between sites

long_diff <- read_csv(here("rda", "data", "long_diff.csv"))

fst_ada_df <- lbl_ada %>% 
  cbind(long_diff) %>% 
  mutate(region1 = case_when(site1 %in% c("BLB", "SHS", "PRS") ~ "Atlantic",
                             site1 %in% c("TCB", "WAB", "APB", "MOB") ~ "Eastern Gulf",
                             site1 %in% c("GAB", "MAB", "SAB", "CCB") ~ "Western Gulf")) %>%
  mutate(region2 = case_when(site2 %in% c("BLB", "SHS", "PRS") ~ "Atlantic",
                             site2 %in% c("TCB", "WAB", "APB", "MOB") ~ "Eastern Gulf",
                             site2 %in% c("GAB", "MAB", "SAB", "CCB") ~ "Western Gulf")) %>%
  mutate(Group = case_when(region1 == "Atlantic" & region2 == "Atlantic"  ~ "Within Atlantic",
                           region1 == "Eastern Gulf" & region2 == "Eastern Gulf"  ~ "Within Eastern Gulf",
                           region1 == "Western Gulf" & region2 == "Western Gulf" ~ "Within Western Gulf", 
                           region1 == "Atlantic" & region2 == "Eastern Gulf" ~ "Atlantic-Eastern Gulf", 
                           region1 == "Atlantic" & region2 == "Western Gulf" ~ "Atlantic-Western Gulf", 
                           region1 == "Eastern Gulf" & region2 == "Western Gulf" ~ "Eastern Gulf-Western Gulf")) %>%
  mutate(Group = ordered(Group, levels = c("Within Atlantic", "Within Eastern Gulf", "Within Western Gulf", "Atlantic-Eastern Gulf", "Atlantic-Western Gulf", "Eastern Gulf-Western Gulf")))

# linear regression by group

## Atlantic-Eastern Gulf

temp <- fst_ada_df %>%
  filter(Group == "Atlantic-Eastern Gulf") %>% 
  drop_na(pwFST)

lm_ada_atl_egom <- lm(temp$long_diff ~ temp$pwFST)

summary(lm_ada_atl_egom)

## Atlantic-Western Gulf

temp <- fst_ada_df %>%
  filter(Group == "Atlantic-Western Gulf") %>% 
  drop_na(pwFST)

lm_ada_atl_wgom <- lm(temp$long_diff ~ temp$pwFST) 

summary(lm_ada_atl_wgom)

## Eastern Gulf-Western Gulf

temp <- fst_ada_df %>%
  filter(Group == "Eastern Gulf-Western Gulf") %>% 
  drop_na(pwFST)

lm_ada_egom_wgom <- lm(temp$long_diff ~ temp$pwFST)

summary(lm_ada_egom_wgom)

## Atlantic-Atlantic

temp <- fst_ada_df %>%
  filter(Group == "Within Atlantic") %>% 
  drop_na(pwFST)

lm_ada_atl_atl <- lm(temp$long_diff ~ temp$pwFST)

summary(lm_ada_atl_atl)

## Eastern Gulf-Eastern Gulf

temp <- fst_ada_df %>%
  filter(Group == "Within Eastern Gulf") %>% 
  drop_na(pwFST)

lm_ada_egom_egom <- lm(temp$long_diff ~ temp$pwFST)

summary(lm_ada_egom_egom)

## Western Gulf-Western Gulf

temp <- fst_ada_df %>%
  filter(Group == "Within Western Gulf") %>% 
  drop_na(pwFST)

lm_ada_wgom_wgom <- lm(temp$long_diff ~ temp$pwFST)

summary(lm_ada_wgom_wgom)

# plot

temp <- fst_ada_df %>% 
  filter(Group == "Within Atlantic" | Group == "Within Eastern Gulf" | Group == "Within Western Gulf") %>% 
  mutate(Group = ordered(Group, levels = c("Within Atlantic", "Within Eastern Gulf", "Within Western Gulf")))

dat_text <- data.frame(
  label = c("ns", "ns", "**"),
  Group   = c("Within Atlantic", "Within Eastern Gulf", "Within Western Gulf"),
  x = c(0.3, 0.3, 0.3),
  y = c(0.0275, 0.0275, 0.0275))

fst_long <- ggplot(data = temp, aes(x = long_diff, y = pwFST), show.legend = F) +
  geom_point(size = 2, alpha = 0.75) +
  geom_smooth(method = lm, size = 1, alpha = 0.5) +
  scale_x_continuous(limits = c(0, 6), breaks = seq(0, 6, 1), labels= seq(0, 6, 1)) +
  scale_y_continuous(limits = c(0, 0.03), breaks = seq(0, 0.03, 0.01), labels= seq(0, 0.03, 0.01)) +
  #facet_wrap(~ Group, ncol = 3) +
  facet_grid(. ~ Group, scales = "fixed", space = "fixed") +
  labs(x = "Absolute Difference in Degrees of Longitude", y = expression(paste("Adaptive Pairwise ", italic(F)[ST]))) +
  ggtitle("B") +
  guides(fill = F, shape = F, color = F) +
  theme_standard +
  theme(legend.text=element_text(size=12), legend.title=element_text(size=12), legend.position = "right", legend.box = "vertical", plot.title = element_text(size = 20), axis.text=element_text(size=12)) +
  geom_text(
  data = dat_text,
  size = c(5, 5, 10),
  color = "red",
  mapping = aes(x = x, y = y, label = label))
        
fst_long

```

### Figure 3

```{r fig.height=12, fig.width=5, message=FALSE, warning=FALSE}

mtdna_gulf / neu_gulf / ada_gulf

ggsave(here("pop_str", "results", "fig3.png"))

ggsave(here("pop_str", "results", "fig3.pdf"))

```

### Figure 4

```{r fig.height=7, fig.width=9.5, message=FALSE, warning=FALSE}

fst_lat / fst_long

ggsave(here("pop_str", "results", "fig4.png"))

ggsave(here("pop_str", "results", "fig4.pdf"))

```
